### 2017-08-13 paperMVVB/sections/acknowledgments.tex
mvvb edits implementation done

\begin{acknowledgements}
	This work was done within the scope of the Coronagraphic German~and~US SolarProbePlus Survey (CGAUSS), funded by the German Aerospace Center (DLR) under grant 50~OL~1601.\\
	The authors thank the Helios and OMNI PIs/teams for creating and making available the solar wind in situ data. The Helios and the OMNI data are supplied by the NASA Space Science Data Coordinated Archive (NSSDCA) and the Space Physics Data Facility (SPDF) at NASA's Goddard Space Flight Center (GSFC).\\
	Additional thanks for maintaining and providing the international sunspot number series goes to the World Data Center -- Sunspot Index and Long-term Solar Observations (WDC-SILSO) at the Solar Influences Data Analysis Center (SIDC), Royal Observatory of Belgium (ROB).\\
	The PSP SPICE kernel was kindly provided by Angelos~Vourlidas.\\
%	We thank the referees for /very important comments and suggestions. /careful review of this paper.

%	facilities/persons which made available the Helios and OMNI data\\
% 	``Ian Richardson and Hilary Cane for making their Interplanetary Coronal Mass Ejection list readily available''\\
% 	``Joseph King and Natalia Papitashvili for creating the combined OMNI data set''\\
% 	``We thank the reviewers for the careful review of this paper''\\
% 	The Authors extend their thanks to the referee for important comments and suggestions.\\
% 	We would like to thank XX for reading drafts of this manuscript...\\
% 	We would like to thank the Helios team and NASA's SPDF for supplying the plasma and magnetic field data (url)\\
\end{acknowledgements}