### 2017-10-12 paperMVVB_referee/referee_report_responses.txt
#3 worked on referee comments; made changes bold face

Referee Report

This paper discusses a method of using OMNI and Helios solar wind data to estimate the solar wind parameters that might be expected during the upcoming Parker Solar Probe mission. It includes an estimate of how these parameters might vary with solar cycle, though this is a relatively small correction, and the likely sunspot number during the mission is anyway highly uncertain. The analysis appears to be thorough, retaining information on the log-normal distribution of the parameters that is used to produce “model” representations of the distributions and their variation during the solar cycle. However, in the end, this information isn’t used in the predictions at PSP, where only median values are considered. It’s also not clear whether this type of analysis has any advantage over say a simple-minded approach where you use the radial dependence of averaged Helios values to propagate averaged 1 AU values, perhaps taken at different times in the solar cycle, down to the PSP orbit. My suspicion is that the results might not be all that different, but the authors might like to address this point. The extrapolated values do appear to be fairly consistent with previous estimates of near Sun parameters, and deviations, e.g. in the speed profile can be plausibly accounted for, so the results appear to be reasonable.

- ...
+ use the distributions for the predictions (add perihelia figures of distributions)
+ state advantage over simple-minded approach (example of simple-minded approach; state how different the results are)
+ maybe stress Helios data bias to cycle minimum (section 4, third paragraph)

Some steps in the procedure could be explained more clearly in the text, as noted below, and the tables are rather cryptic since the meaning of the various parameters is not always clear. If they are fit parameters, it would be helpful to state the function that is being fitted in the table. There are conclusions and summary sections that both, yet again, go through the analysis procedure, which is unnecessary. They could be combined. There are also useful functions that give the radial dependences of the parameters that are only shown in the summary section. They could be included in the main results section.

- ...
+ procedural steps to be explained in detail
+ clarify the meaning of the parameters in tables
+ link the fit equations in table captions
+ combine conclusion and summary sections
+ include final functions in results section

The English is reasonable overall, but there are a number of cases where a long phrase is used as an adjective. Suggestions have been made to rephrase these sentences.

Overall, I anticipate that the paper will be acceptable for publication after the comments in this review are considered.

Specific Points:

Page 2, Results, line 5: Suggest: Since the modeled velocity and temperature values below about 20 Rs appear overestimated in comparison with existing observations, this suggests that PSP will directly measure solar wind acceleration and heating processes below 20 Rs as planned.

- adopted
+ highlighted
+ can be adopted

Introduction, line 8: “Until today” is unnecessary – suggest removing. SOHO is also at L1 and returns solar wind observations from CELIAS. But maybe it’s not necessary to list the spacecraft, especially as other spacecraft not listed contributed earlier observations to OMNI.

- removed
+ highlighted
+ can be removed

- SOHO added
- should we keep it?
+ highlighted
+ add SOHO
+ SOHO/CELIAS not real-time (ca. +3 hours)
+ SOHO/CELIAS does not contribute to OMNI data set

Line 15: Suggest: The in-situ solar wind measurements closest to the Sun to date were made by ...

- adopted
+ highlighted
+ can be adopted

Line 16: Suggest: Helios 1, launched in 1974, reached distances… au. Helios 2, launched two years later, approached the Sun as close as 0.29 au.

- adopted
+ highlighted
+ can be adopted

Page 6, Table 1:

Suggest writing the units together with the parameters in the first column, e.g. “Magnetic field (nT)”, rather than in a table footnote.

- changed
- should we keep it? doesn't it look too packed?
+ highlighted
+ can be adopted

+ do that for all tables

I don’t really understand what the numbers in brackets (“estimated standard deviations”) represent. Why are they much larger than the parameter values and whole numbers?

- added notation explanation
+ highlighted
+ explain

- added notation explanation in all tables
+ explain in all tables??

+ explain also errors??

How are the mean absolute errors and sums of absolute errors defined?

- mean absolute errors unnecessary (they are not referred to in paper), removed from table
+ highlighted

- explained more clearly in text and table caption
+ highlighted
+ explain

Page 7, Figure 1 caption, last sentence, suggest: show zoomed-in views of the high value tails of the distributions.

- adopted
+ highlighted
+ can be adopted

- changed in fig 2 too
+ change in fig 2 too

Line 5 below caption: Supporting reference or other evidence that the extreme values are associated with “CME events” (ICMEs?)?

+ highlighted
+ explaining sentence and reference

Line 7: Is this “overestimation” because [I]CMEs have abnormally low, not high, temperatures?

+ highlighted
+ look into that

Page 8, last paragraph, line 2: Maybe avoid the word "correlated" since it is used below, and there is no correlation here. E.g., … fast solar wind occurs at times when polar coronal holes extend to lower latitudes, a typical feature of the declining phase of the solar cycle as pointed out by..... Therefore... exhibit time lags relative to the SSN maxima.

- changed
+ highlighted
+ change

- adopted
+ highlighted
+ adopt
+ 'time lags' without hyphen!

Page 9, Figure 3: What does the grey shading in the top panels represent? Mention this in the caption.

- explained
+ explain

Three lines below figure caption: It's the amplitudes of the variations in the correlations that decline with increasing lag, not the amplitudes of the correlation coefficients.

- changed
+ highlighted
+ change

Next line: a period [not frequency] of about 11 years.

- adopted
+ highlighted
+ can be adopted

- changed 'lag time' to 'time lag'
+ below, change 'lag time' to 'time lag'

Page 10, Table 2: Similar comments to those for Table 1 apply.

- table caption equation references added
- in section 3 sentence with equation (9) clarified
+ highlighted
+ change

Three lines below table and following: I find this step difficult to follow. It seems that the sunspot numbers are shifted according to the lags, and there is some fit to the data which is then used to produce a model of the particular parameter as a function of SSN, but the text doesn't explain the motivation (apparently to produce a "model" distribution as a function of SSN) procedure very clearly.

- explained more clearly
+ highlighted
+ explain motivation more clearly

- section 3, second sentence: abbreviated 'sunspot number'
- section 3, switched second and third paragraph
- section 3, added a sentence to now second paragraph

Below equation 8: I don't understand what "with the implementation of these relations into the lognormal function" means exactly. How are the fit coefficients obtained?

- phrased this point more clearly
+ highlighted
+ explain

Page 13, Table 3: Give the equation for the fit, to place the parameters in context.

- referenced the fit equation
+ highlighted
+ link the fit equation

- referenced fit equations in all other table captions
+ reference fit equations for all other tables

Third paragraph below table, line 2: “See Figure 7”: Why not explain Figure 7 more fully here, then the coefficients d med, davg, etc. obtained from the fits will make more sense?

- introduced figure properly
- moved second figure reference to end of this paragraph
+ highlighted

Last paragraph of page: “As anticipated…” - especially since the same data are used! Also give the results in the text to make the comparison easier with the values given in this paragraph.

- removed 'As anticipated'
- added exponents into text
+ highlighted

+ is both sufficient?

Page 14, last paragraph, line 6: The curves are still very close together after the crossing, so is this an important feature?

+ highlighted
+ what does that mean? no?

Line 8: But this distance (0.082 AU) is beyond the range shown in the figure, and the data, so is it reliable? I don't understand the comment "therefore the extrapolation of the magnetic field and temperature distribution frequencies to the PSP orbit by applying lognormal functions is limited. What does “limited” mean? Perhaps the comment arises because the mean is larger than the median for a log normal distribution, as noted several pages earlier, and so this crossing indicates that the distribution is no longer log normal. Maybe a few more words of explanation would help the reader understand this point, and the significance of the crossing point.

- explained more clearly
+ highlighted
+ what does limited mean?
+ limited in the sense of a correct distribution shape

+ is it reliable? it is calculated from that data

- clarified the significance of the crossing point
+ highlighted
+ explain significance of crossing point

Line 10: How are the identical exponents chosen? E.g., was emed always used, or eavg, or some other value? This sentence reads as though the values were the same for all four parameters, but presumably it really means that a different exponent was used for each parameter.

- reworded sentence
- added remark in next paragraph
+ highlighted
+ reword to clarify both

Page 15, paragraph 3, line 2: The figure is introduced twice, here and two paragraphs above. Can this be avoided?

- rephrased figure references
+ highlighted
+ aviod this

Line 7, suggest: The increasing high velocity tail with distance comes from.

- adopted
+ highlighted
+ can be adopted

Page 16, line below figure: Not sure what "established" means here. Remove?

- removed
+ highlighted
+ remove

Line 3: What parameter is used for the power law fits? Presumably it's the radial distance, but this could be stated specifically.

- stated specifically and added equation reference
+ highlighted
+ state specifically

Page 17, figure caption: solar wind parameter power exponents for the dependence on radial distance

- adopted
+ highlighted
+ can be adopted

Last sentence of paragraph, suggest: Since this separation is highly variable, and requires significant effort to calculate for an extended time series, we have ignored this aspect in this analysis.

- adopted
+ highlighted
+ adopt reworded

- a few lines before: added 'and' between references 'Bruno et al. (1986) and Balogh et al. (1999)'

Section 6, line 2, suggest: predictions of the SSN during the mission are incorporated into the empirical solar wind model derived in the previous sections

- adopted
+ highlighted
+ can be adopted

Next paragraph, last sentence, suggest: the perihelion distance to be reduced to a minimum of 9.86 Rs. Reverse figures 10 and 11, since figure 11 is referenced first, or just remove the mention of Figure 11 here, which isn't necessary; the statement can be understood without looking at the figure.

- adopted
+ highlighted
+ can be adopted

- switched order of figures 10 and 11
- added top and bottom pointers to both references of now figure 10
+ highlighted
+ manage fig references

Last line of page: How are the ranges of median values for different SSN shown in the figure? I'm guessing that it's the thickness of the shaded regions, but this could be stated specifically to guide the reader in interpreting the figure.

- added remark about that
+ highlighted
+ explain thickness of shaded region

- added further explanation in figure caption

Figure 10 caption, line 1: The models obtained from Helios and OMNI measurements are …

- adopted
+ highlighted
+ can be adopted

- ordered the references Sheeley 1997 and Wang 2000 by year

Line 3: But do the extreme values really occur for min and max SSN? E.g. the highest speeds could be at intermediate SSN? How can a time lag be “noted” in this figure which there is no obvious time dependence?

+ yes, in average, see figure 5

- correct, removed remark about time lag from figure caption
+ remove remark about time lag from caption

- added column to table 2: SSN time lag

Line 4 below figure: “Arguably”: There's no argument here; this just seems to be a suggestion.

- reworded beginning of sentence
+ highlighted
+ reword

Line 5, suggest: mentioned changing shape of the frequency distribution with heliocentric distance

- adopted
+ highlighted
+ can be adopted

Line 6: But the largest difference appears to be at ~1 AU, so isn't there also a problem with the dipole/quadrapole model matching the observed fields at 1 AU?

+ highlighted
+ DQCS model assumes 3.1 nT at 1 au coming from CHs at +-60° latitudes (excluding field from equatorial slow wind belt)

Line 7: “Whereas”: Combine the sentences; whereas isn’t used at the beginning of a sentence. Since the peak speeds are on the declining phase of the cycle, wouldn't the peak predicted speeds occur for some intermediate sunspot number, not at SSN=200?

- combined
+ highlighted
+ combine

- correct, it is a shifted SSN. added value of time lag (at each solar wind parameter)
+ explain

Last line: New sentence at “Thus”.

- adopted
+ highlighted
+ can be adopted

Page 19, line 6: What is the relevance of the comment on assuming a helium abundance? It's not clear how it relates to the comparison of the density profiles.

- reworded
+ highlighted
+ explain

+ is that ok?

Line 10: overestimates

- adopted
+ highlighted
+ can be adopted

Paragraph 2, line 2: There is an abrupt change of topic here, and it's not immediately obvious why short term SSN predictions are relevant. There could be an introductory sentence saying that since the solar wind predictions depend on SSN, they may be refined using predictions of the SSN during the PSP mission.

+ highlighted
+ look into it
+ additional argument is first perihelion

Line 5: Why should the last cycle be used rather than say a previous one? Is this based on expectations that the next cycle will also be small?

+ highlighted
+ explain

Line 8, suggest: the SSN during the closest perihelia, which will commence at the end of 2024 at the likely maximum phase of cycle 25, cannot be predicted at this time. Suggest moving this up to before "For the prediction of the next solar cycle..."

- adopted
+ highlighted
+ can be adopted

- moved up
+ move up

Paragraph 3: How are the error bands in Figures 12 and 13 calculated? E.g., do they use different assumptions for the predicted SSN?

+ answer/explain

Figure 11, lower legend in figure: “Cycle 24” amplitude rather than "its".

- changed
+ fig legend
+ can be adopted

Page 21, line 4, suggest: au, deriving power law fit functions that were used

- adopted
+ highlighted
+ can be adopted

Line 6, suggest: -2025, encompassing the prime mission up to the closest approach of 9.86 Rs.

- adopted
+ highlighted
+ can be adopted

- replaced solar radii with Rs in discussion
+ search text for solar radii and replace with Rs

Paragraph 2, line 3, suggest: Differences from the results of these other studies suggest that below about 20 Rs, PSP may dive into the region where the acceleration and heating of the solar wind is expected to occur.

- adopted with addition
+ highlighted
+ can be adopted, but
+ pay attention to wording

Last line on page, suggest: because the solar wind is believed to be accelerated up to the height of the Alfvenic critical surface, which is predicted to lie on average around 17 Rs, with a range of between 15 Rs at solar minimum and 30 Rs at solar maximum.

- adopted with addition
+ highlighted
+ can be adopted, but
+ check wording

Page 22, paragraph 2, line 5, suggest: that due to solar wind drag, the speeds

- adopted
+ highlighted
+ can be adopted

Summary, paragraph 2: It's not really necessary to go though the procedure yet again; there is already a synopsis in the conclusions. Is it necessary to have both a conclusions and a summary section? They could be combined.

+ highlighted
+ combine both

Equation 12 and following equations: The formulas presented here have not been shown above, but are much easier to interpret than the results above. I'd suggest moving them into the results section above.

+ move equations or explain

Page 23, last but one paragraph: Again, there are points here that have already been said twice above.

+ highlighted
+ remove or merge with above

- added referee acknowledgment